Symbolic Weighted Language Models and\\ Quantitative Parsing over Infinite Alphabets

### 2021-10-10 main.tex
old example 5 in comment
minor

\lydia{We should find a better variable name than j... it does not render well in fractions}
Let's say the top of the stack is a tuple of the call~$\ccall{n}\tstamp{[\tau, \tau+\ell]}$ and a state $q$. Let's now consider a new call transition starting in state $q_{\frac{j}{n}}$, that is the $j$-th state of the $n$ sub-intervals of duration~$\frac{\ell}{n}$.
We thus read a new time-division symbol ~$\ccall{d}$ and compute the weight of the transition from $q_{\frac{j}{n}}$ to $q_{\frac{1}{d}}$ (the first state after the new ~$\ccall{d}$), reading the top of the stack $<\ccall{n}\tstamp{[\tau, \tau+\ell]}, q>$, and pushing $<\ccall{d}\tstamp{\iota}, q_{\frac{j+1}{c}}>$ on top, or:
\(
\weic\bigl(q_{\frac{j}{n}},
           \ccall{n}\tstamp{[\tau, \tau+\ell]}, q,
           \ccall{d}\tstamp{\iota}, q_{\frac{j+1}{c}},
           q_{\frac{1}{d}} \bigr) = \alpha_d
\).
\lydia{is it ok to explain precisely this computation and not the others ? is it redundant with the explaniation before equation (5) ?}
\noindent
Reading the $k$-th musical event $\mu$ in this current sub-interval is computed with:
\(
\weii\bigl(q_{{k}/{d}},
           \ccall{d}\tstamp{[\tau, \tau+\ell]}, q_{\frac{j}{n}},
           \mu\tstamp{\tau + \frac{j\ell}{d}}, q_{\frac{k+1}{d}} \bigr) = \alpha_\mu
\).
Finally, this sub-interval will end after reading a return symbol $\creturn{d}$ and compute:
\(
\weir\bigl(q_{\frac{d}{d}},
           \ccall{d}\tstamp{[\tau, \tau+\ell]}, q_{\frac{i+1}{c}},
           \,\creturn{d}\tstamp{\tau + \ell}, q_{\frac{i+1}{c}} \bigr) = \one
\).
\noindent
We described earlier the special cases where the stack is empty, which applies for example when reading the first measure for which we only push $<\ccall{\mathsf{m}}\tstamp{[0, 1]}, q_{1/1}>$ in the stack:
\(
\weiec\bigl(q_{1/1},
           \ccall{\mathsf{m}}\tstamp{[0, 1]}, q_{1/1},
           q_{1/1} \bigr) = \one
\).
In comparison, any other new measure would compute instead:
\(
\weic\bigl(q_{1/1},
          \ccall{\mathsf{m}}\tstamp{[\tau-1, \tau]}, q_{1/1},
          \ccall{\mathsf{m}}\tstamp{[\tau, \tau+1]}, q_{1/1},
          q_{1/1} \bigr) = \one
\).

@@

@@

\noindent
In order to compute the above function $b_\bot : Q \times Q \to \Semiring$,
we shall consider an auxiliary function $b_\top :  Q \times P \times Q \to \Phic$.
%%%%%% lydaddition %%%%%%
The function $b_\bot$ will allow to compute $b_\top$ recursively in the case where the stack in not empty.
%%%%%% lydaddition %%%%%%
%
Intuitively, $b_\top(q, p, q')$ is a function of $\Phic$,
mapping every $c \in \Deltac$ to
the minimum weight of a computation of~$A$
starting in state $q$ with a non-empty stack
$\gamma' = \<c, p> \, \gamma \in \Gamma^+$,
and ending in state $q'$ with the same stack $\gamma'$,
such that
the computation does not pop the pair $\<c, p>$ at the top of $\gamma'$
(\ie $\gamma'$ is left untouched during the computation).
However, the computation can read $\<c, p>$ at the top of $\gamma'$,
and can also push another pair $\< c', p'> \in \Gamma$ on top of $\gamma'$,
following the third case 
in the definition \eqref{eq:SWVPA-weight} of $\weight_A$ (call symbol).
The pair $\< c', p'>$ can be popped later, during the computation from $q$ to $q'$,
following the fifth case of \eqref{eq:SWVPA-weight} (return symbol).
%However, it cannot apply one of the two last cases (return symbol and empty stack)
%when the current stack is $\gamma$.
%pop symbols in $\gamma$.
%
% Note that having a stack reduced to such a symbol makes impossible the application of the
% two last cases in the definition of $\weight_A$ (return symbol and empty stack).
% However, it is possible to apply the two first cases
% (internal symbol or call symbol, with a push on the top of $\top$).
%
Formally, in order to define $b_\top$, we consider
a fresh stack symbol $\top \notin \Gamma$,   %which does not belong to $\Gamma$,
representing the above untouched stack, and let:
%
\begin{equation}\label{eq:btop}
  b_\top(q, p, q') : c \mapsto \bigoplus_{s\in \Delta^*}
  \textstyle
  \weight_A\bigl(\configup{q}{\< c, p> \stackup \top }, s, \configup{q'}{\<c, p> \stackup \top}\bigr)
\quad
\mathrm{~for~all~} c \in \Deltac
\end{equation}
%
%By definition of $\weight_A$ in \eqref{eq:SWVPA-weight},
%using the symbol $\top$ for the part of the stack below $\<c, p>$
%(\ie the substack $\gamma$ in the above $\gamma' = \<c, p> \, \gamma$)
%ensures that this part is not touched during the computation.
% there cannot be a pop of $\<c, p>$ followed by a push of $\<c, p>$
% because it is not possible to push over $\top$ (push read the top of stack).
%
This ensures in particular that the subword
read during the computation is well parenthesized
(every symbol in $\Deltac$ has a matching symbol in $\Deltar$).

@@

\begin{figure}[ht]
\(
\begin{array}{lcl}
\multicolumn{3}{l}{%
 \mbox{For~every}\  q_0, q_3 \in Q,}\\[1mm] % if $\< q_1, p, q_2> \notin \Q$
%
d_\bot(q_0, q_3) & \opluseq &
  d_\bot(q_0, q_2) \otimes
  \bigoplus_{\Deltai} \weiei(q_2, q_3)\ 
  \mathrm{if~} \< q_0, q_2> \notin \Q \\
%
d_\bot(q_0, q_3) & \opluseq &
  {\bigoplus_{\Deltac}}
  \bigl(
  \weiec(q_0, p, q_1) \otimes
   d_\top(q_1, p, q_2) \otimes
  \bigoplus_{\Deltar} \weir(q_2, p, q_3)\bigr)\ 
  \mathrm{if~} \< q_1, p, q_2> \notin \Q \\
%
d_\bot(q_0, q_3) & \opluseq &
  d_\bot(q_0, q_2) \otimes
  \bigoplus_{\Deltar} \weier(q_2, q_3)\ 
  \mathrm{if~} \< q_0, q_2> \notin \Q \\ 
%
d_\bot(q_0, q_3) & \opluseq &
  d_\bot(q_0, q_2) \otimes d_\bot(q_2, q_3)\  
  \mathrm{if~} \< q_0, q_2>\notin \Q, \< q_2, q_3> \notin \Q \\[-1mm]
\end{array}
\)
\caption{\textbf{update} $d_\bot$ (Algorithm~\ref{algo:Dijkstra}).}
\label{fig:best-update-bottom}
\end{figure}
%

\begin{figure}[ht]
\(
\begin{array}{lcl}
\multicolumn{3}{l}{%
 \mbox{For~every}\  q_0, q_3 \in Q\ \mbox{and}\ p \in P,}\\[1mm] % if $\< q_1, p, q_2> \notin \Q$
d_\top(q_0, p, q_3) & \opluseq &
  d_\top(q_0, p, q_2) \otimes
  \bigoplus_{\Deltai} \weii(q_2, p, q_3)\
  \mathrm{if~} \< q_0, p, q_2> \notin \Q \\
%\multicolumn{3}{l}{\quad
%\mathrm{where~} \weii(q_2, p, q_3)_a \in \Phic
%\mathrm{~is~the~partial~application~}
%x_c \mapsto \weii(q_2, x_c, p, a, q_3)}\\
d_\top(q_0, p, q_3) & \opluseq &
  {\bigoplus^2_{\Deltac}}
  \bigl[ \bigl( \weic(q_0, p, q_1, p) \otimes_2
  d_\top(q_1, p', q_2) \bigr) \otimes_2
  {\bigoplus_{\Deltar}} \weir(q_2, p', q_3) \bigr]\
  \mathrm{if~} \< q_1, p', q_2> \notin \Q \\
%\multicolumn{3}{l}{\quad
%\mathrm{where~} \weic(q_0, p, p', q_1)_{c'} \in \Phic
%\mathrm{~is~the~partial~application~}
%x_c \mapsto \weic(q_0, p, x_c, c', p', q_1)}\\[2pt]
%
d_\top(q_0, p, q_3) & \opluseq &
  d_\top(q_0, p, q_2) \otimes d_\top(q_2, p, q_3)\ 
  \mathrm{if~} \< q_0, p, q_2> \notin \Q, < q_2, p, q_3> \notin \Q \\[-1mm]
\end{array}
\)
%$\weii(q_2, p, q_3)_a \in \Phic$, for $a \in \Deltai$,
%is the partial application
%$x_c \mapsto \weii(q_2, x_c, p, a, q_3)$.
%
%$\weic(q_0, p, p', q_1)_{c'} \in \Phic$, for $c' \in \Deltac$,
%is the partial application
%$x_c \mapsto \weic(q_0, p, x_c, c', p', q_1)$.
%\caption{Update $d_\bot$ with $\<q_1, q_2>$ or $d_\top$ with $\< q_1, p, q_2>$.}
\caption{\textbf{update} $d_\top$ (Algorithm~\ref{algo:Dijkstra}).}
\label{fig:best-update-top}
\end{figure}
%
%\noindent
%** effectively computable by hypothesis that the label theory is effective**
%\florent{REVOIR}
Thanks to the hypothesis that $\bar\Phi$ is effective, 
it is possible to construct during the iteration a witness word
for Proposition~\ref{th:best-search}, 
\ie a word $t \in \Delta^*$ with a minimal weight $A(t)$ \wrt~$\leq_\oplus$.

>>>>

\lydia{We should find a better variable name than j... it does not render well in fractions}
Let's say the top of the stack is a tuple of the call~$\ccall{n}\tstamp{[\tau, \tau+\ell]}$ and a state $q$. Let's now consider a new call transition starting in state $q_{\frac{j}{n}}$, that is the $j$-th state of the $n$ sub-intervals of duration~$\frac{\ell}{n}$.
We thus read a new time-division symbol ~$\ccall{d}$ and compute the weight of the transition from $q_{\frac{j}{n}}$ to $q_{\frac{1}{d}}$ (the first state after the new ~$\ccall{d}$), reading the top of the stack $<\ccall{n}\tstamp{[\tau, \tau+\ell]}, q>$, and pushing $<\ccall{d}\tstamp{\iota}, q_{\frac{j+1}{c}}>$ on top, or:
\(
\weic\bigl(q_{\frac{j}{n}},
           \ccall{n}\tstamp{[\tau, \tau+\ell]}, q,
           \ccall{d}\tstamp{\iota}, q_{\frac{j+1}{c}},
           q_{\frac{1}{d}} \bigr) = \alpha_d
\).
\lydia{is it ok to explain precisely this computation and not the others ? is it redundant with the explaniation before equation (5) ?}
\noindent
Reading the $k$-th musical event $\mu$ in this current sub-interval is computed with:
\(
\weii\bigl(q_{\frac{k}{d}},
           \ccall{d}\tstamp{[\tau, \tau+\ell]}, q_{\frac{j}{n}},
           \mu\tstamp{\tau + \frac{j\ell}{d}}, q_{\frac{k+1}{d}} \bigr) = \alpha_\mu
\).
Finally, this sub-interval will end after reading a return symbol $\creturn{d}$ and compute:
\(
\weir\bigl(q_{\frac{d}{d}},
           \ccall{d}\tstamp{[\tau, \tau+\ell]}, q_{\frac{i+1}{c}},
           \,\creturn{d}\tstamp{\tau + \ell}, q_{\frac{i+1}{c}} \bigr) = \one
\).
\noindent
We described earlier the special cases where the stack is empty, which applies for example when reading the first measure for which we only push $<\ccall{\mathsf{m}}\tstamp{[0, 1]}, q_{\frac{1}{1}}>$ in the stack:
\(
\weiec\bigl(q_{\frac{1}{1}},
           \ccall{\mathsf{m}}\tstamp{[0, 1]}, q_{1/1},
           q_{1/1} \bigr) = \one
\).
In comparison, any other new measure would compute instead:
\(
\weic\bigl(q_{\frac{1}{1}},
          \ccall{\mathsf{m}}\tstamp{[\tau-1, \tau]}, q_{\frac{1}{1}},
          \ccall{\mathsf{m}}\tstamp{[\tau, \tau+1]}, q_{\frac{1}{1}},
          q_{\frac{1}{1}} \bigr) = \one
\).

@@

%\begin{example} \label{ex:SWVPA}
%\reviews{revision ex. needed. see rev. 3) and app. $\weic$, $\weir$, $\weii$ wrong.}
%We consider a \SWVPA over the alphabet of Example~\ref{ex:nested-word} and with $P = Q$
%that expresses a weight related to the music notation,
%or more precisely to its structural complexity.
%Given a set of equivalent representations, we aim at choosing the simpler one.
%
%Let's say the top of the stack is a tuple of the call~$\ccall{n}\tstamp{[\tau, \tau+\ell] and a state $q$. Let's now consider a new call transition starting in state $q_{\frac{j}{n}}$, that is the $j$-th state of the $n$ sub-intervals of duration~$\frac{\ell}{n}$.
%We thus read a new time-division symbol ~$\ccall{d}$ and compute the weight of the transition from q_{\frac{j}{n}} to q_{\frac{1}{d} (the first state after the new ~$\ccall{d}$), reading the top of the stack <\ccall{n}\tstamp{[\tau, \tau+\ell]}, q>, and pushing <\ccall{d}\tstamp{\iota}, q_{\frac{j+1}{c}}> on top, or:
%\(
%\weic\bigl(q_{\frac{j}{n}},
           %\ccall{n}\tstamp{[\tau, \tau+\ell]}, q,
           %\ccall{d}\tstamp{\iota}, q_{\frac{j+1}{c}},
           %q_{\frac{1}{d}} \bigr) = \alpha_d
%%\).
%
%\noindent
%Reading the $k$-th musical event $\mu$ in this current sub-interval is computed with:
%\(
%\weii\bigl(q_{{k}/{d}},
           %\ccall{d}\tstamp{[\tau, \tau+\ell]}, q_{\frac{j}{n}},
           %\mu\tstamp{\tau + \frac{j\ell}{d}}, q_{\frac{k+1}{d}} \bigr) = \alpha_\mu
%\).
%Finally, this sub-interval will end after reading a return symbol \creturn{d} which will compute the transition:
%\(
%\weir\bigl(q_{\frac{d}{d}},
           %\ccall{d}\tstamp{[\tau, \tau+\ell]}, q_{\frac{i+1}{c}},
           %\,\creturn{d}\tstamp{\tau + \ell}, q_{\frac{i+1}{c}} \bigr) = \one
%\).
%
%\noindent
%We described earlier the special cases where the stack is empty, which applies for example when reading the first measure for which we push <\ccall{\mathsf{m}}\tstamp{[0, 1]}, q_{1/1}>, but don't read anything on the stack:
%\(
%\weiec\bigl(q_{1/1},
 %          \ccall{\mathsf{m}}\tstamp{[0, 1]}, q_{1/1},
%           q_{1/1} \bigr) = \one
%\).
%In comparison, any other new measure would compute instead:
%\(
%\weic\bigl(q_{1/1},
%          \ccall{\mathsf{m}}\tstamp{[\tau-1, \tau]}, q_{1/1},
%          \ccall{\mathsf{m}}\tstamp{[\tau, \tau+1]}, q_{1/1},
%          q_{1/1} \bigr) = \one
%\).
%\endex
%\end{example}

@@

\noindent
In order to compute the above function $b_\bot : Q \times Q \to \Semiring$,
we shall consider an auxiliary function $b_\top :  Q \times P \times Q \to \Phic$.
%%%%%% lydaddition %%%%%%
The function $b_\bot$ will allow to compute $b_\top$ recursively in the case where the stack in not empty.
%%%%%% lydaddition %%%%%%
%
Intuitively, $b_\top(q, p, q')$ is a function of $\Phic$,
mapping every $c \in \Deltac$ to
the minimum weight of a computation of~$A$
starting in state $q$, 
with a non-empty stack $\gamma' = \<c, p> \, \gamma \in \Gamma^+$,
and ending in state $q'$ with the same stack $\gamma'$,
such that
the computation does not pop the pair $\<c, p>$ at the top of $\gamma'$
(\ie $\gamma'$ is left untouched during the computation).
However, the computation can read $\<c, p>$ at the top of $\gamma'$,
and can also push another pair $\< c', p'> \in \Gamma$ on top of $\gamma'$,
following the third case 
in the definition \eqref{eq:SWVPA-weight} of $\weight_A$ (call symbol).
The pair $\< c', p'>$ can be popped later, during the computation from $q$ to $q'$,
following the fifth case of \eqref{eq:SWVPA-weight} (return symbol).
%However, it cannot apply one of the two last cases (return symbol and empty stack)
%when the current stack is $\gamma$.
%pop symbols in $\gamma$.
%
% Note that having a stack reduced to such a symbol makes impossible the application of the
% two last cases in the definition of $\weight_A$ (return symbol and empty stack).
% However, it is possible to apply the two first cases
% (internal symbol or call symbol, with a push on the top of $\top$).
%
Formally, in order to define $b_\top$, we consider
a fresh stack symbol $\top \notin \Gamma$,   %which does not belong to $\Gamma$,
representing the above untouched stack, and let:
%
\begin{equation}\label{eq:btop}
  b_\top(q, p, q') : c \mapsto \bigoplus_{s\in \Delta^*}
  \textstyle
  \weight_A\bigl(\configup{q}{\< c, p> \stackup \top }, s, \configup{q'}{\<c, p> \stackup \top}\bigr)
\quad
\mathrm{~for~all~} c \in \Deltac
\end{equation}
%
%By definition of $\weight_A$ in \eqref{eq:SWVPA-weight},
%using the symbol $\top$ for the part of the stack below $\<c, p>$
%(\ie the substack $\gamma$ in the above $\gamma' = \<c, p> \, \gamma$)
%ensures that this part is not touched during the computation.
% there cannot be a pop of $\<c, p>$ followed by a push of $\<c, p>$
% because it is not possible to push over $\top$ (push read the top of stack).
%
This ensures in particular that the subword
read during the computation is well parenthesized
(every symbol in $\Deltac$ has a matching symbol in $\Deltar$).

@@

\begin{figure}[ht]
\(
\begin{array}{lcl}
\multicolumn{3}{l}{%
 \mbox{For~every}\  q_0, q_3 \in Q,}\\[1mm] % if $\< q_1, p, q_2> \notin \Q$
%
d_\bot(q_0, q_3) & \opluseq &
  d_\bot(q_0, q_2) \otimes
  \bigoplus_{\Deltai} \weiei(q_2, q_3)\ 
  \mathrm{if~} \< q_0, q_2> \notin \Q \\
%
d_\bot(q_0, q_3) & \opluseq &
  {\bigoplus_{\Deltac}}
  \bigl(
  \weiec(q_0, p, q_1) \otimes
   d_\top(q_1, p, q_2) \otimes
  \bigoplus^2_{\Deltar} \weir(q_2, p, q_3)\bigr)\ 
  \mathrm{if~} \< q_1, p, q_2> \notin \Q \\
%
d_\bot(q_0, q_3) & \opluseq &
  d_\bot(q_0, q_2) \otimes
  \bigoplus_{\Deltar} \weier(q_2, q_3)\ 
  \mathrm{if~} \< q_0, q_2> \notin \Q \\ 
%
d_\bot(q_0, q_3) & \opluseq &
  d_\bot(q_0, q_2) \otimes d_\bot(q_2, q_3)\  
  \mathrm{if~} \< q_0, q_2>\notin \Q, \< q_2, q_3> \notin \Q \\[-1mm]
\end{array}
\)
\caption{\textbf{update} $d_\bot$ (Algorithm~\ref{algo:Dijkstra}).}
\label{fig:best-update-bottom}
\end{figure}
%

\begin{figure}[ht]
\(
\begin{array}{lcl}
\multicolumn{3}{l}{%
 \mbox{For~every}\  q_0, q_3 \in Q\ \mbox{and}\ p \in P,}\\[1mm] % if $\< q_1, p, q_2> \notin \Q$
d_\top(q_0, p, q_3) & \opluseq &
  d_\top(q_0, p, q_2) \otimes
  \bigoplus_{\Deltai} \weii(q_2, p, q_3)\
  \mathrm{if~} \< q_0, p, q_2> \notin \Q \\
%\multicolumn{3}{l}{\quad
%\mathrm{where~} \weii(q_2, p, q_3)_a \in \Phic
%\mathrm{~is~the~partial~application~}
%x_c \mapsto \weii(q_2, x_c, p, a, q_3)}\\
d_\top(q_0, p, q_3) & \opluseq &
  {\bigoplus^2_{\Deltac}}
  \bigl[ \bigl( \weic(q_0, p, p', q_1) \otimes_2
  d_\top(q_1, p', q_2) \bigr) \otimes_2
  {\bigoplus_{\Deltar}} \weir(q_2, p', q_3) \bigr]\
  \mathrm{if~} \< q_1, p', q_2> \notin \Q \\
%\multicolumn{3}{l}{\quad
%\mathrm{where~} \weic(q_0, p, p', q_1)_{c'} \in \Phic
%\mathrm{~is~the~partial~application~}
%x_c \mapsto \weic(q_0, p, x_c, c', p', q_1)}\\[2pt]
%
d_\top(q_0, p, q_3) & \opluseq &
  d_\top(q_0, p, q_2) \otimes d_\top(q_2, p, q_3)\ 
  \mathrm{if~} \< q_0, p, q_2> \notin \Q, \< q_2, p, q_3> \notin \Q \\[-1mm]
\end{array}
\)
%$\weii(q_2, p, q_3)_a \in \Phic$, for $a \in \Deltai$,
%is the partial application
%$x_c \mapsto \weii(q_2, x_c, p, a, q_3)$.
%
%$\weic(q_0, p, p', q_1)_{c'} \in \Phic$, for $c' \in \Deltac$,
%is the partial application
%$x_c \mapsto \weic(q_0, p, x_c, c', p', q_1)$.
%\caption{Update $d_\bot$ with $\<q_1, q_2>$ or $d_\top$ with $\< q_1, p, q_2>$.}
\caption{\textbf{update} $d_\top$ (Algorithm~\ref{algo:Dijkstra}).}
\label{fig:best-update-top}
\end{figure}
%
%\noindent
%** effectively computable by hypothesis that the label theory is effective**
%\florent{REVOIR}
Thanks to the hypothesis that $\bar\Phi$ is effective, 
it is possible to construct during the iteration a witness word
for Proposition~\ref{th:best-search}, 
\ie a word $t \in \Delta^*$ with a minimal weight $A(t)$ \wrt~$\leq_\oplus$.

### 2021-10-10 proof.tex
minor

Let $<q_0, q_3>$ be the first tuple for which the invariant is false:
$$
For <q_0, q_3> \notin Q, d_{bot} (q_0, q_3) \neq b_{bot} (q_0, q_3)
$$
Since $d_{bot}(q_0, q_2)$ has to be computed to compute $d_{bot}(q_0, q3)$, and because the invariant is true for  $d_{bot}(q_0, q_2)$ we can say:
$$
d_{bot}(q_0, q_2) = b_{bot}(q_0, q_2) \\
d_{bot}(q_0, q_2) \leq b_{bot}(q_0, q_3) \\
\\d_{bot}(q_0, q_2) \leq d_{bot}(q_0, q_3) \\
$$
 However, when $<q_0, q_3>$ was extracted, $<q_0, q_2>$ was still in $Q$. Thus, by definition 
$$
d_{bot}(q_0, q_2) \geq d_{bot}(q_0, q_3) \\
$$
Then the two inequalities are actually equal : 
$$
d_{bot}(q_0, q_2) = b_{bot}(q_0, q_2) = d_{bot}(q_0, q_3) = b_{bot}(q_0, q_3) \\
$$
We find that $d_{bot}(q_0, q_3) = b_{bot}(q_0, q_3)$, so by contradiction, we find that indeed $d_{bot}(q_0, q_3) = b_{bot}(q_0, q_3) iff <q_0, q_3> \notin Q$.

>>>>

Let $<q_0, q_3>$ be the first tuple for which the invariant is false:
$$
For <q_0, q_3> \notin Q, d_{bot} (q_0, q_3) \neq b_{bot} (q_0, q_3)
$$
Since $d_{bot}(q_0, q_2)$ has to be computed to compute $d_{bot}(q_0, q3)$, and because the invariant is true for  $d_{bot}(q_0, q_2)$ we can say:
$$
d_{bot}(q_0, q_2) = b_{bot}(q_0, q_2) \\
d_{bot}(q_0, q_2) \leq b_{bot}(q_0, q_3) \\
\\d_{bot}(q_0, q_2) \leq d_{bot}(q_0, q_3) \\
$$
 However, when $<q_0, q_3>$ was extracted, $<q_0, q_2>$ was still in $Q$. Thus, by definition 
$$
d_{bot}(q_0, q_2) \geq d_{bot}(q_0, q_3) \\
$$
Then the two inequalities are actually equal : 
$$
d_{bot}(q_0, q_2) = b_{bot}(q_0, q_2) = d_{bot}(q_0, q_3) = b_{bot}(q_0, q_3) \\
$$
We find that $d_{bot}(q_0, q_3) = b_{bot}(q_0, q_3)$, so by contradiction, we find that indeed $d_{bot}(q_0, q_3) = b_{bot}(q_0, q_3)$ iff 
$\<q_0, q_3> \notin Q$.
